\section{Overview of the Proof}

The proof of Seymour's decomposition theorem is organized into several largely independent components, each of which is treated separately in this blueprint.

\medskip

\noindent
\textbf{Totally unimodular representations.}  
Chapter~1 develops the theory of totally unimodular matrices, pivoting operations, and standard representations of vector matroids.  
A key result is the characterization of regular matroids as those admitting totally unimodular representations.

\medskip

\noindent
\textbf{Closure under sums.}  
Chapters~2-4 show that regularity is preserved under $1$-, $2$-, and $3$-sums of matroids.  
These chapters contain the technical core of the argument and formalize the matrix constructions underlying each sum operation.

\medskip

\noindent
\textbf{Special matroids.}  
Chapter~5 introduces several distinguished classes of matroids, including graphic matroids, cographic matroids, and the exceptional matroid $R_{10}$.  
The chapter records the relevant definitions and states the regularity results for these classes.

\medskip

\noindent
\textbf{Cographic matroids.}  
Chapter~6 provides a detailed account of the regularity of cographic matroids.  
This chapter serves both to establish the required result and to illustrate the use of duality and orthogonal complements in the context of standard representations.

\medskip

\noindent
\textbf{Conclusion.}  
The final chapter assembles the preceding results into the form required for Seymour's theorem, expressing regular matroids as sums of basic building blocks.
