\chapter{Regularity of 1-Sum}

\begin{definition}
    \label{standardReprSum1}
    \uses{StandardRepr,Matrix.fromBlocks}
    % \lean{}
    \leanok
    Let $R$ be a magma containing zero (we will use $R = \mathbb{Z}_{2}$ and $R = \mathbb{Q}$). Let $B_{\ell} \in R^{X_{\ell} \times Y_{\ell}}$ and $B_{r} \in R^{X_{r} \times Y_{r}}$ be matrices where $X_{\ell}, Y_{\ell}, X_{r}, Y_{r}$ are pairwise disjoint sets. The $1$-sum $B = B_{\ell} \oplus_{1} B_{r}$ of $B_{\ell}$ and $B_{r}$ is
    \[
        B = \begin{bmatrix} B_{\ell} & 0 \\ 0 & B_{r} \end{bmatrix} \in R^{(X_{\ell} \cup X_{r}) \times (Y_{\ell} \cup Y_{r})}.
    \]
\end{definition}

\begin{definition}
    \label{Matroid.IsSum1of}
    \uses{Matroid,StandardRepr,standardReprSum1}
    % \lean{}
    \leanok
    A matroid $M$ is a $1$-sum of matroids $M_{\ell}$ and $M_{r}$ if there exist standard $\mathbb{Z}_{2}$ representation matrices $B_{\ell}$, $B_{r}$, and $B$ (for $M_{\ell}$, $M_{r}$, and $M$, respectively) of the form given in Definition~\ref{standardReprSum1}.
\end{definition}

\begin{lemma}
    \label{Matrix.det_fromBlocks_zero}
    \uses{Matrix.det,Matrix.fromBlocks}
    % \lean{}
    \leanok
    Let $A$ be a square matrix of the form $A = \begin{bmatrix} A_{11} & A_{12} \\ 0 & A_{22} \end{bmatrix}$. Then $\det A = \det A_{11} \cdot \det A_{22}$.
\end{lemma}

\begin{proof}
    \uses{Matrix.det,Matrix.fromBlocks}
    \leanok
    This result is proved in Mathlib.
\end{proof}

\begin{lemma}
    \label{Matrix.fromBlocks_isTotallyUnimodular}
    \uses{standardReprSum1,Matrix.IsTotallyUnimodular}
    % \lean{}
    \leanok
    Let $B_{\ell}$ and $B_{r}$ from Definition~\ref{standardReprSum1} be TU matrices (over $\mathbb{Q}$). Then $B = B_{\ell} \oplus_{1} B_{r}$ is TU.
\end{lemma}

\begin{proof}
    \uses{standardReprSum1,Matrix.IsTotallyUnimodular,Matrix.det_fromBlocks_zero}
    \leanok
    We prove that $B$ is TU by Definition~\ref{Matrix.IsTotallyUnimodular}. To this end, let $T$ be a square submatrix of $B$. Our goal is to show that $\det T \in \{0, \pm 1\}$.

    Let $T_{\ell}$ and $T_{r}$ denote the submatrices in the intersection of $T$ with $B_{\ell}$ and $B_{r}$, respectively. Then $T$ has the form
    \[
        T = \begin{bmatrix} T_{\ell} & 0 \\ 0 & T_{r} \end{bmatrix}.
    \]

    First, suppose that $T_{\ell}$ and $T_{r}$ are square. Then $\det T = \det T_{\ell} \cdot \det T_{r}$ by Lemma~\ref{Matrix.det_fromBlocks_zero}. Moreover, $\det T_{\ell}, \det T_{r} \in \{0, \pm 1\}$, since $T_{\ell}$ and $T_{r}$ are square submatrices of TU matrices $B_{\ell}$ and $B_{r}$, respectively. Thus, $\det T \in \{0, \pm 1\}$, as desired.

    Without loss of generality we may assume that $T_{\ell}$ has fewer rows than columns. Otherwise we can transpose all matrices and use the same proof, since TUness and determinants are preserved under transposition. Thus, $T$ can be represented in the form
    \[
        T = \begin{bmatrix} T_{11} & T_{12} \\ 0 & T_{22} \end{bmatrix},
    \]
    where $T_{11}$ contains $T_{\ell}$ and some zero rows, $T_{22}$ is a submatrix of $T_{r}$, and $T_{12}$ contains the rest of the rows of $T_{r}$ (not contained in $T_{22}$) and some zero rows. By Lemma~\ref{Matrix.det_fromBlocks_zero}, we have $\det T = \det T_{11} \cdot \det T_{22}$. Since $T_{11}$ contains at least one zero row, $\det T_{11} = 0$. Thus, $\det T = 0 \in \{0, \pm 1\}$, as desired.
\end{proof}

\begin{theorem}
    \label{Matroid.IsSum1of.isRegular}
    \uses{Matroid.IsSum1of,Matroid.IsRegular}
    % \lean{}
    \leanok
    Let $M$ be a $1$-sum of regular matroids $M_{\ell}$ and $M_{r}$. Then $M$ is also regular.
\end{theorem}

\begin{proof}
    \uses{StandardRepr,Matroid.IsSum1of,Matroid.IsRegular,StandardRepr.toMatroid_isRegular_iff_hasTuSigning,Matrix.fromBlocks_isTotallyUnimodular,Matrix.IsTuSigningOf}
    \leanok
    Let $B_{\ell}$, $B_{r}$, and $B$ be standard $\mathbb{Z}_{2}$ representation matrices from Definition~\ref{Matroid.IsSum1of}. Since $M_{\ell}$ and $M_{r}$ are regular, by Lemma~\ref{StandardRepr.toMatroid_isRegular_iff_hasTuSigning}, $B_{\ell}$ and $B_{r}$ have TU signings $B_{\ell}'$ and $B_{r}'$, respectively. Then $B' = B_{\ell}' \oplus_{1} B_{r}'$ is a TU signing of $B$. Indeed, $B'$ is TU by Lemma~\ref{Matrix.fromBlocks_isTotallyUnimodular}, and a direct calculation shows that $B'$ is a signing of $B$. Thus, $M$ is regular by Lemma~\ref{StandardRepr.toMatroid_isRegular_iff_hasTuSigning}.
\end{proof}
