\subsection{Totally Unimodular Matrices}

In our work, the notion of totally unimodular matrices is key to defining regular matroids. Before stating what a totally unimodular matrix is, let us review how matrices and submatrices are implemented in Mathlib. Matrices with rows indexed by \texttt{m}, columns indexed by \texttt{n}, and entries of type \texttt{α} are represented by \texttt{Matrix m n α}, implemented as a (curried) binary function \leaninline{m → n → α}. This allows the user to access the elements of matrix $A$ with \texttt{A i j}. Similarly, \texttt{Matrix.submatrix} is defined so that \texttt{(A.submatrix f g) i j = A (f i) (g j)} holds. For example, if
\[
    A = \begin{bmatrix}
       1 & 2 & 3 \\
       4 & 5 & 6 \\
       7 & 8 & 9
    \end{bmatrix}
\]
and \texttt{f = ![0]} and \texttt{g = ![2, 2, 0, 0]}, then
\[
    \texttt{A.submatrix f g} =
    \begin{bmatrix}
       3 & 3 & 1 & 1
    \end{bmatrix}
\]
typed as a matrix (not a vector). In particular, note that \texttt{Matrix.submatrix} may repeat and reorder rows and columns.

Now, a matrix $A$ over a commutative ring $R$ is \emph{totally unimodular} if every finite square submatrix of $A$ (not necessarily contiguous, with no row or column taken twice) has determinant in $\{-1, 0, 1\}$. Mathlib implements this definition as follows:
\begin{leancode}
def Matrix.IsTotallyUnimodular {m n R : Type*}
    [CommRing R] (A : Matrix m n R) :
    Prop :=
  ∀ k : ℕ, ∀ f : Fin k → m, ∀ g : Fin k → n,
    f.Injective → g.Injective →
      (A.submatrix f g).det ∈
        Set.range SignType.cast
\end{leancode}
Here, \texttt{SignType} is an inductive type with three values \texttt{zero}, \texttt{neg}, and \texttt{pos}, and \texttt{SignType.cast} maps them to \texttt{(0 : R)}, $\texttt{(-1 : R)}$, and $\texttt{(1 : R)}$, respectively.

Also note that the indexing functions \texttt{f} and \texttt{g} are required to be injective in the definition, but this condition can be lifted. Indeed, lemma \texttt{Matrix.isTotallyUnimodular\_iff} shows that one can equivalently check the determinants of all finite square submatrices, not just ones with no repeated rows or columns. This characterization may be easier to work with than the original definition, as it lifts the additional condition on \texttt{f} and \texttt{g}.

Finally, note that the determinant is computed over $R$, so for certain commutative rings, all matrices are trivially totally unimodular, for example, if $R = \mathbb{Z}_{3}$.
